---
title: "Contrasts"
output: html_document
---

First, let's download and visualize a dataset for ANOVA.

```{r}
library("broom")
library("dplyr")
library("ggplot2")
library("gmodels")
library("readr")
library("reshape2")
theme_set(theme_bw())
```

```{r}
etch <- read_csv("https://uwmadison.box.com/shared/static/3ltmo89ea0xowsh1386x9fk58qc51ned.txt")
etch$Power <- as.factor(etch$Power)
etch
```

Before, we can an anova on the difference between power levels, and found
(unsurprisingly) that there was a big difference in etch rate between power
levels.

```{r}
m_etch <- etch %>%
  melt(measure.vars = paste0("Ob", 1:5), value.name = "etch_rate") %>%
  arrange(Power)

fit <- lm(etch_rate ~ Power, data = m_etch)
aov_fit <- aov(fit)
summary(aov_fit)
```

Let's say we're interested in the contrast between power levels of 160 and 180.
We have to encode this as a contrast object, and then pass it to the
`fit.contrast` function.

```{r}
contrast <- c(1, -1, 0, 0)
fit.contrast(aov_fit, "Power", contrast)
```

We can also do several contrasts at a time.

```{r}
contrasts <- matrix(
    c(1, -1, 0, 0,
      1, 1, -1, -1),
    nrow = 2, byrow = TRUE
  )
fit.contrast(aov_fit, "Power", contrasts)
```

And you can get confidence intervals by passing in the `conf.int` parameter.

```{r}
fit.contrast(aov_fit, "Power", contrasts, conf.int = 0.95)
```