---
title: "Diagnostics and Power"
output: html_notebook
---

```{r}
library("ggplot2")
library("dplyr")
library("reshape2")
theme_set(theme_bw())
```

# Normal Probability Plots

First, let's get a sense of how normal probability plots work. Note that the
axes are reversed compared to the definition in the textbook.

```{r}
x <- rnorm(1000)
tau <- 1.8
x[x > tau] <- runif(sum(x > tau), 1, 1.4) * x[x > tau]
hist(x, 50)
```
```{r}
qqnorm(x)
abline(a = 0, b = 1, col = "red")
```
# Power Analysis

We'll write some code to do a power analysis in the two sample t-test. We'll see
how varying the underlying signal strength and the collected sample size
influence the test results.

First, let's write a small function to simulate data from a set of normal
distributions.

```{r}
simulate_data <- function(true_means, n, sigma=1) {
  X <- matrix(nrow = n, ncol = length(true_means))
  for (i in seq_len(n)) {
    X[i, ] <- rnorm(length(true_means), true_means, sigma)
  }

  melt(X, varnames = c("index", "group"))
}
```

Next, we'll simulate many runs across a range of signal strengths.

```{r}
alpha <- 0.05
N <- 1000 # sample size per group
mu_seq <- seq(0, .5, 0.05) # power levels
B <- 200 # number of reps per level
test_result <- list()

# loop over signal strengths levels
for (i in seq_along(mu_seq)) {
  print(mu_seq[i])
  
  # many reps for signal level
  test_result[[i]] <- vector(length = B)
  for (j in seq_len(B)) {
    x <- simulate_data(c(0, mu_seq[i]), N)
    test_result[[i]][[j]] <- t.test(x$value[x$group == 1], x$value[x$group == 2])$p.value
  }
  
  # save a plot
  p <- ggplot(x) +
    geom_histogram(aes(x = value, fill = as.factor(group)), position = "identity", alpha = 0.7) +
    scale_fill_brewer(palette = "Set2") +
    ggtitle(sprintf("mu[1] = %s", mu_seq[i]))
  print(p)
}

names(test_result) <- mu_seq
mtest_result <- melt(test_result)
```

Now that we've simulated the data and run all the tests, we can look at the average power.

```{r}
colnames(mtest_result) <- c("p.value", "mu")
ggplot(mtest_result) +
  geom_hline(aes(yintercept = 0.05), col = "red") +
  geom_jitter(aes(x = mu, y = p.value)) +
  scale_y_log10()

ggplot(mtest_result %>%
         group_by(mu) %>%
         summarise(power = mean(p.value < 0.05))) +
  geom_point(aes(x = mu, y = power))
```

