---
title: "Difference in Means"
output: html_notebook
---

We'll download the cement data that's described in section 2.4.

```{r}
library("broom")
library("ggplot2")
library("readr")
library("reshape2")
cement <- read_csv("https://uwmadison.box.com/shared/static/sjpn7dmlvcyp5wvo9orn8euuk24yvkq9.txt")
cement
```

For visualization, it's going to be better to have the data in "tall" or "tidy"
format.

```{r}
mcement <- melt(cement, measure.vars = c("Modified", "Unmodified"), variable.name = "formulation", value.name = "strength")
head(mcement)
```

```{r}
theme_set(theme_bw())
ggplot(mcement) +
  geom_point(aes(x = formulation, y = strength))
```

Now that we've made a preliminary exploratory view, let's run a t-test. The 
null hypothesis is that there's no difference in the average strength of the 
modified and unmodified cements.

```{r}
t_result <- t.test(cement$Modified, cement$Unmodified, var.equal = TRUE)
t_result
```

We might want a simple way to extract different statistics from the t-test
result. This is made easier through the `tidy` function in the `broom` package.

```{r}
tidy(t_result)
```

It turns out that the $t$-test result is equivalent to running a linear
regression with just one predictor.

```{r}
fit <- lm(strength ~ formulation, data = mcement)
summary(fit)
```


```{r}
t.test(cement$Modified, cement$Unmodified, var.equal = FALSE)
```