---
title: "Factorial Design Intro"
output: html_notebook
---

Let's look at the structure of a factorial design experiments. We won't fit any
models, but we'll build some intuition for how the data are usually structured
and how you can go about plotting it.

```{r}
library("dplyr")
library("ggplot2")
library("readr")
library("reshape2")
theme_set(theme_bw())
```

We'll download the battery dataset, which asked how manufacturing material and
temperature both can influence the life of a battery.

```{r}
battery <- read_table2("https://uwmadison.box.com/shared/static/vmxs2wcsdxkdjujp85nw5kvk83xz4gl9.txt") %>%
  mutate(Material = as.factor(Material))
```

We'll make two plots. The first plots the two factors as two axes and shows the
response as the color at each coordinate. The second uses the principle of small
multiples, and shows the relationship between temperature and life, for each
type of material.

```{r}
ggplot(battery) +
  geom_point(
    aes(x = Material, y = Temperature, col = Life),
    position = position_jitter(w = 0.1, h = 1)
  ) +
  scale_color_viridis_c()

ggplot(battery) +
  geom_point(aes(x = Temperature, y = Life)) +
  facet_wrap(~ Material)
```

```{r}
ggplot(battery) +
  geom_point(aes(x = Material, y = Life)) +
  facet_wrap(~Temperature)
```

