---
title: "Latin Squares"
output: html_notebook
---

As usual, we'll load some libraries.
```{r}
library("readr")
library("agricolae")
library("ggplot2")
library("dplyr")
theme_set(theme_bw())
```

Before looking at a data example, let's check out the `agricolae` library, which
makes it easy to generate different Latin Squares.
```{r}
trt <- LETTERS[1:10]
latin_square <- design.lsd(trt)
head(latin_square$book)
ggplot(latin_square$book) + 
  geom_tile(aes(x = row, y = col, fill = trt))
  #scale_fill_brewer(palette = "Set1")
```

Now that we see how we would generate a latin square in theory, let's analyze
the results of a study that used a latin square in its design.
```{r}
rocket <- read_table2("https://uwmadison.box.com/shared/static/ac68766l3zcjog1ldrzki3lis74bbd71.txt") %>%
  mutate(
    Batch = as.factor(Batch),
    Operator = as.factor(Operator),
    Formulation = as.factor(Formulation)
  )
```

Compare the table below with table 4.9 in the book. It makes clear that the
study was a latin square design. We can also make a visualization of the design,
because reading tables of numbers is hard.
```{r}
rocket %>%
  dcast(Batch ~ Operator, value = "Formulation")

ggplot(rocket) +
  geom_tile(aes(x = Operator, y = Batch, fill = Formulation)) +
  scale_fill_brewer(palette = "Set1")
```

Given this design, we can fit the model using a linear model. Here,
$$\alpha_{i}, \tau_{j}$$, and $$\beta_{k}$$ are the batch, formulation, and
operator, respectively. We'll use the shorthand `y ~ .` to refer to the model
using all the other variables in the data frame as inputs.
```{r}
#fit <- lm(BurningRate ~ Batch + Operator + Formulation, data = rocket) # gives exact same result
fit <- lm(BurningRate ~ ., data = rocket)
summary(fit)
```

Compare the ANOVA table below with Table 4.12.
```{r}
anova(fit)
```