---
title: "Multiple Comparisons"
output: html_document
---

First, let's download and visualize a dataset for ANOVA.

```{r}
library("DescTools")
library("broom")
library("dplyr")
library("ggplot2")
library("gmodels")
library("readr")
library("reshape2")
theme_set(theme_bw())
```

This code just makes the same anova fit from before.

```{r}
etch <- read_csv("https://uwmadison.box.com/shared/static/3ltmo89ea0xowsh1386x9fk58qc51ned.txt")
etch$Power <- as.factor(etch$Power)
m_etch <- etch %>%
  melt(measure.vars = paste0("Ob", 1:5), value.name = "etch_rate") %>%
  arrange(Power)

fit <- lm(etch_rate ~ Power, data = m_etch)
aov_fit <- aov(fit)
summary(aov_fit)
```

Let's define a few different contrasts, and then use Scheffe's method to adjust
the confidence intervals from an uncorrected fit. We'll use the `PostHocTest` in
the `DescTools` package, which unfortunately has a slightly different syntax
than `fit.contrasts` in `gmodels`.

```{r}
contrasts <- matrix(
    c(1, -1, 0, 0,
      1, 1, -1, -1,
      0, 0, 1, -1),
    nrow = 3, byrow = TRUE
  )
rownames(contrasts) <- c("160 - 180", "160,180 - 200,220", "200 - 220")
fitc <- fit.contrast(aov_fit, "Power", contrasts, conf.int = 0.95)
scheffe_fitc <- PostHocTest(aov_fit, method = "scheffe", conf.level = 0.95, contrast = t(contrasts))
```



```{r}
fitc[, c("lower CI", "upper CI")]
scheffe_fitc$Power[, c("lwr.ci", "upr.ci")]
```

If we just care about all pairwise comparisons, we could just use Tukey's HSD.

```{r}
TukeyHSD(aov_fit)
```

If you don't actually care about experimentwise error, you can use Fisher's LSD.

```{r}
PostHocTest(aov_fit, method = "lsd", conf.level = 0.95)
```