---
title: "Nonparametric ANOVA"
output: html_notebook
---

```{r}
library("readr")
library("reshape2")
library("dplyr")
library("broom")
library("ggplot2")
theme_set(theme_bw())
```

This is the same etch rate data we used in ordinary ANOVA.

```{r}
etch_wide <- read_csv("https://uwmadison.box.com/shared/static/3ltmo89ea0xowsh1386x9fk58qc51ned.txt") %>%
  mutate(Power = as.factor(Power))
etch <- melt(etch_wide, id.vars = "Power", value.name = "etch_rate")
```

The `R` function that performs nonparametric ANOVA is called `kruskal.test`. The
result is significant, but note the slight loss in power, compared to the
earlier parametric ANOVA.

```{r}
kruskal.test(etch_rate ~ Power, data = etch)
```

We can get an intuitive feel for how this test works by looking at a plot of
ranks againt power level.

```{r}
etch <- etch %>%
  mutate(rank = rank(etch_rate))
etch

ggplot(etch) +
  geom_point(aes(x = Power, y = rank))
```

To see why this nonparametric approach might be handy, consider the sensitivity of the standard ANOVA to outliers.

```{r}
sim_data_ <- matrix(rnorm(500), ncol = 4)
sim_data_[c(1, 2), 1] <- runif(2, 100, 250)
sim_data <- melt(sim_data_, varnames = c("sample", "group"))

ggplot(sim_data) +
  geom_point( aes(x = group, y = value))
```

The ANOVA is very sensitive to this outlier, but the nonparametric ANOVA doesn't
think there's much of a difference between these groups (because the one outlier
is converted to just a rank).

```{r}
anova(lm(value ~ group, data = sim_data))
kruskal.test(value ~ group, data = sim_data)
```