---
title: "Randomized Complete Block Designs"
output: html_notebook
---

Let's load some libraries. `lme4` should catch your eye -- it's the library that
will always come up wheenver we talk about random effects.

```{r}
library("dplyr")
library("ggplot2")
library("lme4")
library("readr")
library("reshape2")
theme_set(theme_bw())
```

We'll download the graft data again.
```{r}
graft <- read_table2("https://uwmadison.box.com/shared/static/0ciblk4z2f3k6zizbj4plg3q33w1d0n6.txt") %>%
  melt(id.vars = "Pressure", variable.name = "batch", value.name = "yield") %>%
  mutate(Pressure = as.factor(Pressure))
```

Let's fit a model with a random effect on the blocks (the batch number). The
notation for random effects is the same as it was from before. Compare the
results below with Table 4.6. They are the same maximum likelihood fit, there
are just some differences in how they the results are reported. For example, the
$$R^2$$ is only reported in the JMP output, and the output from `lmer` displays
intervals for $\sigma_{\tau}$ rather than $\sigma_{\tau}^{2}$.

```{r}
fit <- lmer(yield ~ Pressure + (1 | batch), data = graft)
summary(fit)
anova(fit)
confint(fit)
```