---
title: "Randomized Complete Block Designs"
output: html_notebook
---

Let's load some libraries.

```{r}
library("DescTools")
library("dplyr")
library("ggplot2")
library("readr")
library("reshape2")
theme_set(theme_bw())
```

We'll download the vascular graft data, in which we wanted to study the effect
of pressure on yield, but controlling for possible batch effects.

```{r}
graft <- read_table2("https://uwmadison.box.com/shared/static/0ciblk4z2f3k6zizbj4plg3q33w1d0n6.txt") %>%
  melt(id.vars = "Pressure", variable.name = "batch", value.name = "yield") %>%
  mutate(Pressure = as.factor(Pressure))
```

Before doing any analysis, let's make some plots to see if there was any strong
batch effect. We can also make a first pass at whether pressure might be related
to yield in any detectable way.

```{r}
ggplot(graft) +
  geom_point( aes(x = batch, y = yield, col = Pressure)) +
  scale_color_brewer(palette = "Set2")

ggplot(graft) +
  geom_text( aes(x = Pressure, y = yield, col = Pressure, label = batch)) +
  scale_color_brewer(palette = "Set2")
```
Now, let's quantify the relationship between pressure and yield, controlling for
additive batch effects.

```{r}
fit <- lm(yield ~ Pressure + batch, graft)
summary(fit)
```

Compare the ANOVA table, confidence intervals, and point predictions from below with the output of Figure 4.2.

```{r}
anova(fit)
confint(fit)
predict(fit)
```

To check the additivity assumption, let's plot the residuals against the batch.
It seems mostly okay (no strong patterns in residuals across batches), except
possibly for batch 6. The residuals there seem too narrow, and it might be worth
following-up with whoever ran the experiments on that batch.

```{r}
resids <- resid(fit)
qqnorm(resids)
qqline(resids, col = "red")

graft <- graft %>%
  mutate(resid = resids)

ggplot(graft) +
  geom_point(aes(x = batch, y = resid, col = Pressure))
```

If we want, we can also search for some interesting contrasts (correcting for
multiple comparisons), using the same functions as in the multiple comparisons
videos.

```{r}
contrasts <- matrix(
  c(1, 1, -1, -1,
    1, 0, 0, -1),
  nrow = 2, byrow=TRUE
)

PostHocTest(aov(fit), method = "scheffe", contrasts = t(contrasts)) # scheffe's method
TukeyHSD(aov(fit))  # tukey's test
``` 
