---
title: "STAT 424 HW2"
subtitle: Fall 2020
output:
  pdf_document: default
  word_document: default
---

## Instructions

1. Submit your solution on Canvas as a pdf file before *September 25 at 11:59pm*.
2. Use R for all computation, and include your code. 
3. We recommend you start with the template `hw2.Rmd` file on Canvas.

## Rubric

* [30 points] One problem below will be graded for,
  - [15 points] Correctness. Are all parts answered correctly?
  - [15 points] Clarity. Are all answers justified? For questions with a coding
  component, is code provided, and is it readable?
* [20 = 4 * 5 points] Four problems below will be graded on completeness.
* Total: 50 points

```{r setup, include=FALSE}
library("tibble")
knitr::opts_chunk$set(echo = TRUE)
```

## Problems

**1. [DM 2.17, modified]** The viscosity of a liquid detergent is supposed to
average 800 centistokes^[Whatever that means.] at 25&deg;. A random sample of 14
batches of detergent is collected, and the average viscosity is 819. Suppose
that we know that viscosity is normally distributed and has a standard deviation
of $\sigma = 24$ centistokes.

  a. State the hypothesis that should be tested.
  b. Draw a picture summarizing the data in the problem. 
  b. Test these hypotheses using $\alpha = 0.05$. What are your conclusions?
  c. What is the $p$-value for the test?
  d. Find a 95\% confidence interval for the mean.

**2. [DM 2.27]** Photoresist is a light sensitive material applied to semiconductor
wafers so that the circuit pattern can be imaged onto the wafer. After
application, the coated wafers are baked to remove the solvent in the
photoresist mixture and to harden the resist. Here are measurements of the
photoresist thickness (in kA) for eight wafers baked at two different
temperatures. Assume that all of the runs were made in random order.

```{r}
photoresist <- tibble(
  "sample_id" = as.factor(seq(8)),
  "95" = c(11.156, 7.089, 8.097, 11.539, 11.291, 10.799, 6.467, 8.315),
  "100" = c(5.261, 6.748, 7.461, 7.015, 8.133, 7.438, 3.772, 7.964),
)
```

  a. Is there evidence to support the claim that the higher baking temperature
  results in wafers with a lower mean photoresist thickness? Use $\alpha = 0.05$.
  b. What is the $p$-value for the test conducted in part (a)?
  c. Find a 95\% confidence interval on the difference in means. Provide a
  practical interpretation of this interval.
  d. Make a visualization to assist in interpreting the results from this
  experiment.
  e. Check the assumption of normality of the photoresist thickness.
  f. Find the power of this test for detecting an actual difference in means of 2.5 kA.
  g. What sample size would be necessary to detect an actual difference in means
  of 1.5kA with a power of at least 0.9?

**3. [DM 3.26, modified]** Four chemists are asked to determine the percentage of
methyl alcohol in a certain chemical compound. Each chemist makes three
determinations, and the results are the following:

```{r}
chemists <- tibble(
  chemist = c(1, 1, 1, 2, 2, 2, 3, 3, 3, 4, 4, 4),
  ma = c(84.89, 84.24, 84.18, 85.15, 85.13, 84.88, 84.79, 84.48, 85.16, 84.20, 
         84.10, 84.55)
)
```

  a. Make a plot describing the measurements from the four chemists. Describe
  what you see.
  b. Do the chemists differ significantly? Use $\alpha = 0.05$.
  c. Analyze the residuals from this experiment.
  d. If chemist 2 is a new employee, construct a meaningful set of orthogonal
  contrasts that might have been useful at the start of hte experiment.
  
**4. [EDMES 1.8]** The computer science department of a university wishes to test
whether there are differences among three programming texts with respect to how
long it takes a student completing the text to write a program in C++ language.
Eight students from an introductory class were randomly assigned each of the
texts, asked to complete the steps outlined in the text, and then given a
program to write. The results, in minutes, are shown below.  Conduct a
one-factor ANOVA to determine if hte texts are equally effective.


```{r}
books <- tibble(
 book_id = c(rep("1", 8), rep("2", 8), rep("3", 8)),
 minutes = c(11, 9, 12, 13, 15, 12, 13, 12, 14, 12, 14, 13, 15, 15, 14, 16, 12, 
             8, 10, 12, 13, 10, 14, 11)
)
```

**5. [DM 3.23]** The response time in millisceonds was determined for three
different types of circuits that could be used in an automatic valve shutoff
mechanism^[Sounds important, probably should pay attention.]. The results from a
completely randomized experiment are shown in the following table.

```{r}
shutoff <- tibble(
  "1" = c(9, 12, 10, 9, 14),
  "2" = c(20, 22, 23, 17, 30),
  "3" = c(6, 5, 7, 16, 7)
)
```

  a. Plot the data. You may find it useful to reshape the data into tidy
  format.
  b. Test the hypothesis that the three circuit types have the same response
  time. Use $\alpha = 0.01$.
  c. Use Tukey's test to compare pairs of treatment means. Use $\alpha = 0.01$.
  d. Overlay a 95\% confidence interval for the group means on the plot that you
  made in part (a).

## Feedback

a. How much time did you spend on this homework?
b. Which problem did you find most valuable?

## References

[DAE] Montgomery, Douglas C. Design and analysis of experiments. John wiley &
sons, 2017.

[EDMES] Berger, P. D., & Maurer, R. E. (2002). Experimental Design with
Application in Management, Engineering, and the Sciences. CA (USA): Wadsworth
Group Belmont. Accessed for free at
https://search.library.wisc.edu/catalog/991013677228502128.
