---
title: "Interpreting effects in $2 ^ 3$ Designs"
subtitle: "Week 8 [4]"
output:
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
header-includes:
- \usepackage[fontsize=14pt]{scrextend}
---
Reading: 6.3, 6.4

```{r, include=FALSE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

knitr::opts_chunk$set(message = FALSE, warning = FALSE, size = "small")
```

```{r}
library("dplyr")
library("ggplot2")
library("readr")
library("tufte")
theme_set(theme_bw())
```

* As before, the $SS$ terms can be obtained by squaring the contrasts and dividing by the number of data points

	* Can build associated ANOVA tables
	
* We can view the data through a regression. The full model would be

$$y_{i} = \beta_0 + \sum_{k = 1}^{3} \beta_k x_{ik} + \sum_{\text{pairs } k, k^\prime} \beta_{k k^\prime} x_{ik}x_{ik^{\prime}} + \epsilon_{i}$$
though we will often be interested in whether a submodel (which discards some of
the main or interaction effects) can do as well.
	
* To compare a full and a submodel, use the relative sums of squares,

$$R^2 = \frac{SS_{Model}}{SS_{Total}} = 1 - \frac{SS_{E}}{SS_{Total}}$$

(draw a picture that compares the numerator and denominator), in low and high noise situations)

* Instead of trying to understand the entire model’s importance, we might want
to understand the importance of specific terms
* For this, it’s useful to have an uncertainty estimate

_Example calculation_: Variance estimate for effect of A.
\begin{align*}
\text{Var}\left(Effect A\right) &= \text{Var}\left(\frac{1}{2^{K - 1} n}\left(a - b - c + ab + ...\right)\right) \\
&= \left(\frac{1}{2^{K - 1} n}\right)^2\text{Var}\left(a - b - ac + ab + ...\right)
\end{align*}

But remember that $a$ refers to the sum of all samples at corner $a$, and likewise for $b$, $ac$, etc.,
\begin{align*}
\text{Var}\left(a - b - ac + ab + ...\right) &= \text{Var}\left(\sum_{\text{corner } a}y_{i} - \sum_{\text{corner }b}y_{i} - \sum_{\text{corner }ac}y_{i} + ...\right) \\
&= \sum_{\text{corner } a}\text{Var}\left(y_i\right) + \sum_{\text{corner }b}\text{Var}\left(y_i\right) + ...  \\
&= 2^K n \sigma^2
\end{align*}

so at the end of the day, we get
\begin{align*}
\text{Var}\left(Effect A\right) &= \frac{\sigma^2}{2^{K - 2}n}
\end{align*}

and we can estimate $\sigma^2$ by the error sum fo squares $S^2$. From these
variance estimates, we can build confidence intervals that summarize all the
effects.

## Code Example throughout: [Plasma Etching]

## Generalization: $2^K$ designs

Everything we’ve spoken about can be generalized to the case of arbitrary
numbers of factors. For example, the table notation can be used to get effect
estimate for interaction ABCD listed before equation 6.22 in the book, and the
sum of squares remain just the normalized square of the contrasts.