---
title: "Foldover in $2^{K - p}$ Designs"
output:
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
link-citations: yes
header-includes:
- \usepackage[fontsize=14pt]{scrextend}
---
Reading: 8.6, 8.7

```{r, include=FALSE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

knitr::opts_chunk$set(message = FALSE, warning = FALSE, size = "small", echo = FALSE, fig.margin = TRUE, fig.height = 5, cache = TRUE)
```

```{r}
library("dplyr")
library("ggplot2")
library("readr")
library("EBImage")
library("reshape2")
library("ggplot2")
library("tufte")
theme_set(theme_bw() + theme(legend.position = "bottom"))
```
The main drawback of fractional designs is that we can end up with aliased
effects. However, there are specific ways to follow-up initial experiments in a
way that dealiases these effects.

* Best case: The hierarchical end hereditary principles allow strong effects to
be deduced using only a $\frac{1}{2^p}$ fraction of the samples needed for the
full design.

* Worst case: Strong effects remain aliased, and a follow-up dealiasing
experiment is required.

We’ll discuss two dealiasing strategies, full-foldover and single-factor
foldover. They are both particularly relevant in resolution III designs, where
main effects can be confounded with order-2 interactions.

## Full Foldover

* Use case: We want to delias *main effects* for *all factors* under study

	* Assume that there aren’t any interaction effects with order > 2
	
* Idea: For a second fractional factorial run, reverse the signs of all factors

What does this do? Let’s consider an example. The design below corresponds to a
$2^{7 - 3}_{III}$ fractional factorial experiment (`Seq == 1`) that has then
been folded over (`Seq == 2`) by reversing the signs of factors. You can tell
this from the plot below.

Let’s consider the alias structure for the original fractional factorial,
ignoring all interactions of order 3 and higher.

Parsing this matrix, the effects derived from alias groups are

\begin{align*}
[A] = A + BD + CE + FG \\
[B] = B + AD + CF + EG \\
[C] = C + AE + BF + DG \\
[D] = D + AB + CG + EF \\
[E] = E + AC + BG + DF \\
[F] = F + BC + BG + DE \\
[G] = G + CD + BE + AF
\end{align*}

Now, suppose we reversed the signs of all the factors. What happens to the alias
groups? The signs for the second order interactions flip! The resulting effect
estimates are

\begin{align*}
[A]^{fold} = A - BD - CE - FG \\
[B]^{fold} = B - AD - CF - EG \\
[C]^{fold} = C - AE - BF - DG \\
[D]^{fold} = D - AB - CG - EF \\
[E]^{fold} = E - AC - BG - DF \\
[F]^{fold} = F - BC - BG - DE \\
[G]^{fold} = G - CD - BE - AF
\end{align*}

The punchline is that we can now estimate the main effects without any aliasing,

\begin{align*}
A = \frac{1}{2}\left(\left[A\right] + \left[A\right]^{fold}\right) \\
B = \frac{1}{2}\left(\left[B\right] + \left[B\right]^{fold}\right) \\
C = \frac{1}{2}\left(\left[C\right] + \left[C\right]^{fold}\right) \\
D = \frac{1}{2}\left(\left[D\right] + \left[D\right]^{fold}\right) \\
E = \frac{1}{2}\left(\left[E\right] + \left[E\right]^{fold}\right) \\
F = \frac{1}{2}\left(\left[F\right] + \left[F\right]^{fold}\right) \\
G = \frac{1}{2}\left(\left[G\right] + \left[G\right]^{fold}\right) \\
\end{align*}

This is in fact a general principle for dealising main effects from second order
interactions. When you flip the signs of all factors in an original fractional
factorial design, you will get a cancellation of second-order terms when you add
pairs of effect estimates.

## Data Example

Let’s use these ideas to study the effects in the eye data experiment.

## Single-Factor Foldover

* Use case: We want to dealias *main and interaction effects* associated with a
*single factor* in the study

* Idea: For a second fractional factorial run, reverse the signs of just the
factor of interest.

The mechanics at work here are similar to those in the full foldover. As before,
let’s focus attention on the $2^{7 - 3}_{III}$.  Remember that the effect
estimates were,

\begin{align*}
[A] = A + BD + CE + FG \\
[B] = B + AD + CF + EG \\
[C] = C + AE + BF + DG \\
[D] = D + AB + CG + EF \\
[E] = E + AC + BG + DF \\
[F] = F + BC + BG + DE \\
[G] = G + CD + BE + AF
\end{align*}

Suppose we flip the sign of factor $D$ in the follow-up run. The new effect
estimates would be,

\begin{align*}
[A]^{fold} = A - BD + CE + FG \\
[B]^{fold} = B - AD + CF + EG \\
[C]^{fold} = C + AE + BF + DG \\
[D]^{fold} = -D + AB + CG + EF \\
[E]^{fold} = E + AC + BG - DF \\
[F]^{fold} = F + BC + BG - DE \\
[G]^{fold} = G - CD + BE + AF
\end{align*}

You can see this in the plot — all the signs from terms involving $D$ are flipped.

In particular, notice that we can estimate the main effect of $D$ using

\begin{align*}
D = \frac{1}{2}\left(\left[D\right] - \left[D\right]^{fold}\right)
\end{align*}

and interactions involving $D$ using, for example,

\begin{align*}
AD = \frac{1}{2}\left(\left[B\right] - \left[B\right]^{fold}\right).
\end{align*}
