---
title: "Interpreting Effects in $2^2$ Designs"
subtitle: "Week 8 [1]"
output:
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
header-includes:
- \usepackage[fontsize=14pt]{scrextend}
---
Reading: 6.2

```{r, include=FALSE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

knitr::opts_chunk$set(message = FALSE, warning = FALSE, size = "small")
```

```{r}
library("dplyr")
library("ggplot2")
library("readr")
library("tufte")
theme_set(theme_bw())
```

We will often want to know whether estimated main or interaction effects are
significant. We can use ANOVA, though be wary when $n$ is small.

* The numerators in the effect estimate expressions will be called _contrasts_
for the estimated effect
* For example, the contrast for the effect of $A$ is $ab + a - b -
\left(1\right)$.
* The associated sum of squares is
$$
\frac{1}{2^2 n}\left(Contrast\right)^2
$$
for example,
$$
SS_A = \frac{1}{2^2 n}\left[ab + a - b - (1)\right]^2
$$
The associated ANOVA decomposition is
$$
SS_T = SS_A + SS_B + SS_AB + SS_E
$$
and since the factors all have two levels, the associated DF’s for the main and
interaction terms are all 1. The df of $SS_T$ is $n 2^2 - 1$ (number of
samples minus one). Taking the ratio between main and interaction $SS$ terms and
$SS_E$ gives the basis for F-statistics in the ANOVA table.

## Regression View

Another way of summarizing the $2^2$ model is to write a regression,
$$
y = \beta_0 + \beta_1 x_1 + \beta_2 x_2 + \epsilon
$$
where the $x_k$’s take on one of two values, depending on whether or not factor $k$ is active.

* We’ve only included main effects. An interaction would be added via $\beta_{12} x_{1}x_{2}$

* If the factors are binary (on vs. off), we can use a binary encoding. 

* What if they are actually continuous?

	* Use a _natural coding_, which converts low and high levels to ${-1, 1}$.
	
	* Model will apply to all values in interval $[-1, 1]$.
	
(show an associated response surface)
