---
title: "Projection and Blocking in $2^{K - p}$ Designs"
output:
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
link-citations: yes
header-includes:
- \usepackage[fontsize=14pt]{scrextend}
---

Reading: 8.3, 8.4

```{r, include=FALSE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

knitr::opts_chunk$set(message = FALSE, warning = FALSE, size = "small", echo = FALSE, fig.margin = TRUE, fig.height = 5, cache = TRUE)
```

```{r}
library("dplyr")
library("ggplot2")
library("readr")
library("EBImage")
library("reshape2")
library("ggplot2")
library("tufte")
theme_set(theme_bw() + theme(legend.position = "bottom"))
```

To be fluent in working with fractional factorial designs, it’s worth knowing
how to project and block them. They provide useful alternative perspectives of
the designs we have already constructed.

## Projection

* Projection makes it possible to turn a fractional factorial screening
experiment into a full factorial experiment

* Effectively, you delete the columns of your design matrix involving the
factors outside the ones you want to project down to

* Depending on what design you project, a $2^{K - p}$ design will become
either (1) a full factorial design or (2) a $2^{K - r}$ fractional factorial
design, where $r < p$.

* Beware that conclusions will be somewhat tentative, because effects in the
full factorial could be aliased with high-order effects from factors that have
been projected out

Warm-up: Projecting a $2 ^ {4 - 1}$ Design

Consider the $2 ^ {4 - 1}$ design with defining relation $I = ABCD$.

| A | B | C | D |
| --- | --- | --- | --- |
| - | - | - | - |
| + | - | - | + |
| - | + | - | + |
| - | - | + | + |
| + | + | - | - |
| + | - | + | - | 
| - | + | + | - |
|+ | + | + | + |

What happens when you project out A?

| B | C | D |
| --- | --- | --- |
| - | - | - |
| - | - | + |
| + | - | + |
| - | + | + |
| + | - | - |
| - | + | - | 
| + | + | - |
| + | + | + |

You can check that this has all corners of the cube B-C-D, so it’s a full $2^{3}$ design.

2^5 - 1 from 2 ^ 6 - 2 

Sample example from before, with A projected out.

* The associated complete defining relations are $I = ABCE = BCDF = ADEF$.

| B | C| D | E = ABC | F = BCD |
| --- | --- | --- | --- | --- | --- |
 | - | - | - | - | - |
 | - | - | - | + | - |
 | + | - | - | + | + |
 | - | + | - | + | + |
 | - | - | + | - | + |
 | + | - | - | - | + |
 | - | + | - | - | + |
 | - | - | + | + |  + |
 | + | + | - | - | - |
 | + | - | + | + | - |
 | - | + | + | + | - |
 | + | + | - | + | - |
 | + | - | + | - | - |
 | - | + | + | - | - |
 | + | + | + | - | + |
 | + | + | + | + | + |

What about projecting out $E$ and $F$? That’s a full $2 ^ 4$ factorial with one
replicate. It’s just our original basic design.

What about projecting out $A$ and $B$? Still a $2 ^ 4$ factorial with one replicate.

| C| D | E = ABC | F = BCD |
| --- | --- | --- | --- | --- | --- |
 | - | - | - | - |
 | - | - | + | - |
 | - | - | + | + |
 | + | - | + | + |
 | - | + | - | + |
 | - | - | - | + |
 | + | - | - | + |
 | - | + | + |  + |
 | + | - | - | - |
 | - | + | + | - |
 | + | + | + | - |
 | + | - | + | - |
 | - | + | - | - |
 | + | + | - | - |
 | + | + | - | + |
 | + | + | + | + |
 
What about projecting out A and E? This is different! Now we have a $2 ^ {4 -
1}$ factorial with 2 replicates. We can see that only half of the cornners are
present, but each one that is present is present twice.

| B | C | D | F = BCD |
| --- | --- | --- | --- | --- | --- |
 | - | - | - | - |
 | - | - | - | - |
 | + | - | - | + |
 | - | + | - | + |
 | - | - | + | + |
 | + | - | - | + |
 | - | + | - | + |
 | - | - | + |  + |
 | + | + | - | - |
 | + | - | + | - |
 | - | + | + | - |
 | + | + | - | - |
 | + | - | + | - |
 | - | + | + | - |
 | + | + | + | + |
 | + | + | + | + |
 
This ends up being a general rule. If none of the defining words contain all the
factors that you’re projecting out, then you’re left with a full factorial
design; otherwise you get a fractional factorial.

* E.g., AE is part of the first defining relation, so we got a fractional
factorial.

Exercise: Try projecting onto A, B, C. Is it a fractional or full factorial? How
many replicates each? Could you have determined this, just by looking at the
defining relations?

* A design is said to have _projectivity_ $v$ is collapsing to any subset of
$v$ factors will result in a full factorial design. The example above has
projectivity 3.

## Blocking

Let’s say we’ve subsetted to a set of runs corresponding to a fractional
factorial design. But, we still have to break the runs into batches, we have too
many runs to be able to reasonably run them in one block. How can we
appropriately block a fractional factorial design?

* The typical solution is to choose a high-order effect and alias it with the
batches. This divides the runs into two blocks, those where the effect is + and
those where it is -.

* Ensure that the alias group doesn’t contain any low-order effects of interest
(e.g., we don’t want the blocks to be confounded with any main effects)

* We effectively sacrifice the alias group that is confounded with the block.

Example: For our $2 ^{6 - 2}$ design above, we can confound $ACD$ with the
block, to avoid aliasing any main effects with the block effect. This let us
divide our runs into two blocks of size 8.

| A | B | C	| D | E = ABC | F = BCD | Block = ACD |
| --- | --- | --- | --- | --- | --- | --- |
| - | - | - | - | - | - | - (B1) | 
| + | - | - | - | + | - | + (B2) |
| - | + | - | - | + | + | - (B1) |
| - | - | + | - | + | + | + (B2) |
| - | - | - | + | - | + | + (B2) |
| + | + | - | - | - | + | + (B2) |
| + | - | + | - | - | + | - (B1) |
| + | - | - | + | + |  + | - (B1) |
| - | + | + | - | - | - | + (B2) |
| - | + | - | + | + | - | + (B2) |
| - | - | + | + | + | - | - (B1) |
| + | + | + | - | + | - | - (B1) |
| + | + | - | + | - | - | - (B1) |
| + | - | + | + | - | - | + (B2) |
| - | + | + | + | - | + | - (B1) |
| + | + | + | + | + | + | + (B2)|
