---
title: "Nested Designs"
output:
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
link-citations: yes
header-includes:
- \usepackage[fontsize=14pt]{scrextend}
---

Readings: 14.1

```{r, include=FALSE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

knitr::opts_chunk$set(message = FALSE, warning = FALSE, size = "small", echo = FALSE, fig.margin = TRUE, fig.height = 5, cache = TRUE)
```

```{r}
library("dplyr")
library("ggplot2")
library("readr")
library("EBImage")
library("reshape2")
library("tufte")
theme_set(theme_bw() + theme(legend.position = "bottom"))
```

Factorial designs cover the situation where effects from multiple sources are
crossed with one another. In many situations, however, effects are nested within
each other.

* You rely on several suppliers to provide raw material and are interested in
whether some suppliers provide consistently better material. The material
arrives in batches, and there is variation across batches. The batch effect is
*nested within* the supplier effect.

* The hardness of different alloys are to be compared. Several ingots are built
from each alloy, and repeated measures are drawn from each ingot. The ingot
effect is nested within the alloy effect.

## Model

We will consider the two-stage design here. The data are imagined to be drawn
according to

\begin{align*}
y_{ijk} &= \mu + \alpha_{i} + \beta_{j\left(i\right)} + \epsilon_{ijk}.
\end{align*}

* ~$\alpha_i$~ is the parent-effect (e.g., supplier effect)

* ~$\beta_{j\left(i\right)}$~ is the nested-effect associated with parent ~$i$~
(e.g., effect of the ~$j^{th}$~ batch within the ~$i^{th}$~ supplier)

* ~$\epsilon_{ijk} \sim N\left(0, \sigma^{2}\right)$~ is independent noise.

The key difference from the usual factorial model is that the nested effects
~$\beta_{j\left(i\right)}$~ are not the same across different parents ~$i$~.

## Inference

_Fixed-Effects_

First, imagine treating all the effects as fixed. In this case, there are two typical hypotheses of interest.

* Null parent-effect,

\begin{align*}
&H_{0}: \alpha_{i} = 0 \text{ for all } i
&H_{1}: \alpha_{i} \neq 0 \text{ for at least one } i
\end{align*}
	
* Null child-effect,

\begin{align*}
&H_{0}: \beta_{j\left(i\right)} = 0 \text{ for all } i,j
&H_{1}: \beta_{j\left(i\right)} \neq 0 \text{ for at least one } i,j
\end{align*}

In either case, a test is performed using a sum-of-squares decomposition,
\begin{align*}
SS_{T} &= SS_A + SS_{B\left(A\right)} + SS_{E}
\end{align*}

which is similar in structure to those we have seen before, except we now have a nested effect term,
\begin{align*}
\sum_{i = 1}^{a}\sum_{j = 1}^{b} \left(\bar{y}_{ij\cdot} - \bar{y}_{i\cdot\cdot}\right)^{2}.
\end{align*}

which measures how much the ~$j^{th}$~ effect within parent ~$i$~ varies from
the average in that group. The distribution of the associated mean-squares can
then be used to perform an ANOVA. Let’s see an implementation, based on the
supplier-materials example above. (example 14.1)

_Random Effects_
Rather than caring about specific parent or nested effects, we may simply want
to know the typical variation due to either factor. For example, we may not care
about the effect of the 2nd batch in the 3rd supplier, but we may be curious
about the typical size of batch-to-batch variation. In this case, it makes sense
to use random effects. We can use a random effect for the nested factors,

or for both the nested and parent factors,

However, it makes no sense to treat the parent as random, but its child as
fixed.


## Diagnostics

Nested designs are assessed in the same way as other linear models. We should
check normality of the errors and look for potential patterns in the residuals.