---
title: "$2^2$ Factorial Designs"
subtitle: "Week 8 [1]"
output:
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
header-includes:
- \usepackage[fontsize=14pt]{scrextend}
---


```{r, include=FALSE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

knitr::opts_chunk$set(message = FALSE, warning = FALSE, size = "small")
```

```{r}
library("dplyr")
library("ggplot2")
library("readr")
library("tufte")
theme_set(theme_bw())
```

A special case of general factorial design is the $2^K$ design. This arises when
there are $K$ factors, but only 2 levels for each factor. We assume $n$ samples
at each configuration of factor levels.

* With only two levels for each factor, these experiments aren’t useful for
teasing out subtle variations across levels of a factor
* They *are* useful for determining which of a large number of factors might be
worth investigating further (factor screening)

The simplest case of a $2^K$ design is when the number of factors $K = 2$. The experimental design can be represented by a square.

## Notation

It will be handy to define ways of indexing corners of the square. One approach
is to write + or - for whether we are at a low or high level for that factor.
Alternatively, we can represent the corner by all the letters that are at high
levels.

| A | B | label |
|---|---|---------|
| - | - | (1)     |
| + | - | a       |
| - | + | b       |
| + | + | ab      |

We abuse notation and write $a$ to represent the *total* of the response values
at the corner +-, rather than just the index of that corner.

## Estimating effects

Since there are only two levels for each factor, there are transparent formulas
for estimating main and interaction effects.

* The main effect for A summaries the average change in the response when A is
activated. It is defined as
\begin{align}
A &= \frac{1}{2n}\left(\left(ab +a\right) - \left(b + (1)\right)\right)
\end{align}

which is the average of the responses on the edge where A is active minus the
average when A is inactive. The definition for B is analogous.

* The interaction effect measures the degree to which the effect of A changes
depending on whether or not B is active. It is defined as
\begin{align}
AB &= \frac{1}{2n}\left[\left(ab - b\right) + \left(a - \left(1\right)\right)\right]
\end{align}

* The role of A and B is symmetric — you could read the interaction as how the
effect of B changes depending on whether A is active.

* All these effect estimates can be summarized by our tabular notation

|label | effect A | effect B | effect AB |
| --- |---|---|---------|
| (1) | - | - |  + |
| a | + | - | - |
| b | - | + | - |
| ab | + | + | + |
