---
title: "$2 ^ 3$ Factorial Design"
subtitle: "Week 8 [3]"
output:
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
header-includes:
- \usepackage[fontsize=14pt]{scrextend}
---
Reading: 6.3

```{r, include=FALSE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

knitr::opts_chunk$set(message = FALSE, warning = FALSE, size = "small")
```

```{r}
library("dplyr")
library("ggplot2")
library("readr")
library("tufte")
theme_set(theme_bw())
```

* How can we generalize the $2^2$ analysis so that we can do studies that
inspect many factors simultaneously (all at two levels each)?

* We’ll analyze the 3 factor case (the $2^3$ design) first, but in search of
general patterns

* For the $2^3$ design, we have 8 factor configurations

	* Visualize as corners of a cube
	
	* Call the third factor $C$.

## Effect Estimates

Our table notation can be extended to deal with all 8 corners of the cube.

| A | B | C | label |
|---|---|--- | ---------|
| - | - | - | (1) |
| + | - | - | a       |
| - | + | - | b       |
| - | - | + | c       |
| + | + | - | ab      |
| + | - | + | ac      |
| - | + | + | bc      |
| + | + | + | abc   |

* The main effect estimates can be made by subtracting the + from the - corners
 	* Average of when factor is on vs. off
	* Average of one face of the cube minus the average of the opposite

For example,
$$A = \frac{1}{2^2 n}\left[\left(a + ab + ac + abc\right) - \left(\left(1\right) + b + c + bc\right)\right]$$

To estimate interactions, look at how average effects change when you condition.

| B | Average A Effect | 
| --- | ------| 
| + | $\frac{1}{2^2 n}\left[\left(abc - bc\right) + \left(ab - b\right)\right]$ |
| - | $\frac{1}{2^2 n}\left[\left(ac - c\right) + \left(a - \left(1\right)\right)\right]$ |
| Difference | $\frac{1}{2^2 n}\left[abc - bc + ab - b - ac + c - a + \left(1\right)\right]$ |

(Draw a picture for this)

And notice that the associated contrast can be obtained by multiplying the
columns in the table above.

| A | B | C | AB | label |
| --- | --- | --- | --- | ------| 
| - | - | - | + | (1) |
| + | - | - | - | a |
| - | + | - | - | b |
| - | - | + | + | c |
| + | + | - | + | ab |
| + | - | + | - | ac |
| - | + | + | - |  bc |
| + | + | + | + | abc |

The three-way interaction is defined as the change in two-way interactions
across the two values for the third variable. We won’t prove why this works, but
you can use it as a device for avoiding having to memorize everything.