---
title: "Designs for Response Surfaces"
output:
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
link-citations: yes
header-includes:
- \usepackage[fontsize=14pt]{scrextend}
---

Reading: 11.4.1 - 11.4.2

```{r, include=FALSE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

knitr::opts_chunk$set(message = FALSE, warning = FALSE, size = "small", echo = FALSE, fig.margin = TRUE, fig.height = 5, cache = TRUE)
```

```{r}
library("dplyr")
library("ggplot2")
library("readr")
library("EBImage")
library("reshape2")
library("tufte")
theme_set(theme_bw() + theme(legend.position = "bottom"))
```
We have been fitting surfaces to find some configuration of factors that optimizes a response. Sometimes we found the optimum just from one experiment, but more often, we ran a series of experiments, gradually refining our understanding of the response surface.

*For any of these experiments, what designs should we use?*

In theory, any of the designs we have discussed in this course could be used,
but we will prefer those which give us an accurate estimate of the response
surface (especially around stationary points) using as few samples as possible.

## Designs for First-Order Models

In first-order models, we are interested in linear effects associated with each
factor. Reasonable choices in this setting are,

* $2^{K}$ factorial designs

* $2^{K - p}$ fractional factorial designs

* $2^{K}$ and $2^{K - p}$ designs that have been augmented with center
points. These center points allow estimation of measurement error $\sigma^2$
even when there is no replication.

* Simplex designs. In two dimensions, this looks like the corners of an
equilateral triangle (as opposed to a square, which is the $2^{2}$ design.)

## Designs for Second-Order Models

For second-order models, we need to estimation nonlinear and interaction
effects, which means we need more intensive sampling. The most common choice is
the,

* Central Composite Design (CCD): As discussed in Section 6.8, this is a full
factorial $2^{K}$ design that has been supplemented by center and axial
points.

though some alternatives are,

* Box-Behnken design: The associated samples are all on edges of the cube (none
at vertices or in the interior).

* Equiradial design: This generalizes the simplex idea above. Instead of
sampling at the corners of an equilateral triangle, we can sample at corners or
regular polygons (+ center points)

## Variance of Response Surfaces

At any point $x$, it’s possible to estimate the variance of the surface at
that point, denoted $V\left(\hat{y}\left(x\right)\right)$.

* For any response surface, we can also imagine an accompanying variance
surface.

* Away from factor configurations that we’ve sampled, we expect the variance to increase.

If there are only 2 factors, this surface can be directly visualized. 

* But in higher-dimensions, this is impossible.

* In that case, use a _variance dispersion graph_ to visualize the variance as a function of the distance from the center point of the design.

	* Specifically, plot the maximum, minimum, and average variances along rings that emanate from 0
	
* If the variances are constant along each ring, the design is called _rotatable_