---
title: "Optimizing Multiple Responses"
output:
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
link-citations: yes
header-includes:
- \usepackage[fontsize=14pt]{scrextend}
---

Reading: 11.3

```{r, include=FALSE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

knitr::opts_chunk$set(message = FALSE, warning = FALSE, size = "small", echo = FALSE, fig.margin = TRUE, fig.height = 5, cache = TRUE)
```

```{r}
library("dplyr")
library("ggplot2")
library("readr")
library("EBImage")
library("reshape2")
library("tufte")
theme_set(theme_bw() + theme(legend.position = "bottom"))
```

In more complicated systems, we may want to optimize several objectives
simultaneously. More often than not, the goals will be at odds with one another.
For example, we may want to maximize yield while controlling costs. How can we
use response surface methods when we have several competing objectives?

## Overlaying Contours

* The most direct approach is to simply fit several response surfaces.

* Visually inspect results to find factor configurations with desirable values across each response.

## Constrained Optimization

Whenever visual inspection is challenging, mathematical formalizations can offer
support. One idea is to frame the multiple response surface problem as a
constrained optimization.

* Define acceptable ranges for responses  $y_{2}\left(x\right), \dots,
y_{R}\left(x\right)$ that are important, but not our main focus.

* Optimize the response $y_{1}\left(x\right)$ that’s our main focus.

Formally, we look for a configuration of factors $x_{\ast}$ that solves the
optimization
\begin{align*}
\underset{x}{\text{maximize}}\medspace &y_{1}\left(x\right) \\
\text{subject to }\medspace &\left(y_{2}\left(x\right), \dots, y_{R}\left(x\right)\right) \in \mathcal{C}
\end{align*}

where $C$ is the predefined acceptable region for the secondary responses.

## Desirability Functions

The main downside of the constrained optimization approach is that it forces us
to choose one response to prioritize over all others. What if we care about each
response more or less equally?

One idea is to optimize a sort of (geometric) averaged response,
\begin{align*}
\underset{x}{\text{maximize}}\medspace \left[\prod_{r = 1}^{R} y_{r}\left(x\right)\right]^{\frac{1}{R}}
\end{align*}

The issue with this idea is that it treats all responses exactly equally. What
if we want to maximize some, but minimize others? What if we want some to be
near some target value?

The solution is to use _desirability functions_. A few are plotted below. You
can adjust their shape so that the $r^{th}$ desirability function is large for
the values of $y_{r}\left(x\right)$ which are good (sloping down when you want
to minimize, sloping up when you want to maximize).

Then, instead of optimizing the raw averaged response, we optimize the averaged
response after first passing through the desirability functions,
\begin{align*}
\underset{x}{\text{maximize}} \medspace \left[\prod_{r = 1}^{R} d_{r}\left(y_{r}\left(x\right)\right)\right]^{\frac{1}{R}}
\end{align*}
