---
title: "Unreplicated $2^K$ Designs"
subtitle: "Week 9 [1]"
output:
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
header-includes:
- \usepackage[fontsize=14pt]{scrextend}
---

Reading: 6.5


```{r, include=FALSE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

knitr::opts_chunk$set(message = FALSE, warning = FALSE, size = "small", echo = FALSE)
```

```{r}
library("dplyr")
library("ggplot2")
library("readr")
library("tufte")
theme_set(theme_bw())
```


* Sometimes, people will only take only one measurement per factor configuration

	* When $K$ is large, replication can increase the number of samples needed
	substantially
	
	*  E.g., changing n from 1 to 2 when K = 5 means 32 more runs
	
* For factor screening experiments, typically want to allocate samples to
understanding new factors, rather than replicating known configurations

Without replicates to gauge measurement noise, we may encounter two opposite problems,

* Missing a true effect
* Spurious effects

Let’s see how these problems arise, and discuss some solutions.

## Missing True Effects

* If the effect is weak, then if only nearby levels are tested, the effect will
be easy to miss

* Fix: We can space out the levels at which we test each factor

(example figure when not spaced out)

## Spurious Effects

* If there are no replicates, then we can perfectly interpolate the data

* Leaves us with no degrees-of-freedom for estimating $\sigma^2$

	* $\sigma^2$ is needed to perform ANOVA
	
* Fix: Pool together higher-order interactions

	* High-order interactions are typically rare (sparsity of effects principle)
	
	* Can use pooled interaction estimates to obtain $\sigma^2$
	
	* Pooling can only make testing more conservative
	
(Show lm with and without estimates of sigma^2)

## Design Projections

* Related to pooling, sometimes it is clear that a certain factor has no bearing
on the response

* In this case, we may consider removing that factor and all the interaction
terms that include it

* By collapsing the factor, we double the number of replicates

(show cube collapsed into a square)

## Visualing Effects

Graphical methods provide useful summaries for evaluating interactions in the $2^K$ model.

* Daniel Plots

  * If none of the factors had any influence on the response, then the effects
  would all be normally distributed around 0.
  
  * Idea: make a normal probability plot of the effects, and look for those which
  deviate from identity line. These are likely real effects.

* Lenth Plots

  * An alternative is to simply plot the effect sizes directly
  
  * $1.5 \times \text{median}\left(\text{Contrast}_j\right)$ is often used as a
  cutoff point, for the same reason that $1.5IQR$ is often used to detect
  outliers.