---
title: "Variations on Nested Designs"
output:
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
link-citations: yes
header-includes:
- \usepackage[fontsize=14pt]{scrextend}
---

Readings: 14.2, 14.3

```{r, include=FALSE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

knitr::opts_chunk$set(message = FALSE, warning = FALSE, size = "small", echo = FALSE, fig.margin = TRUE, fig.height = 5, cache = TRUE)
```

```{r}
library("dplyr")
library("ggplot2")
library("readr")
library("EBImage")
library("reshape2")
library("tufte")
theme_set(theme_bw() + theme(legend.position = "bottom"))
```

Nested designs are versatile; they can be applied to many levels of nesting and
in conjunction with factors.

## General Nested Designs

Almost the same methodology used in 2-level nested designs carries over to more
general levels. Graphically, the tree of effects grows deeper. Since there are
more parameters at deeper levels of nesting, it is often a good idea to fit
those parameters using random effects.

The example below analyzes a depth-3 nested design using a random effect for the
depth 2 and 3 factors.

## Simultaneous Nesting and Factors

Nested terms can be included in standard factorial models. Consider the
following setting (Example 14.2),

* An industrial engineering is optimizing the assembly time of an electronic
device.

* There are different possible layouts and fixtures of the circuit boards which
may influence the time to assembly.

* The experiment must be run across factories. The operators who do the actual
assembly will differ depending on layout.

The operator effect is nested within the layout effect. It should also be
treated as a random effect, because we want to understand variation across all
possible operators, when choosing a particular layout and fixture design.
Therefore, a reasonable model is

\begin{align*}
y_{ijkl} &= \mu + \alpha_{i} + \beta_{j} + \tau_{k\left(j\right)} + \left(\alpha\beta\right)_{ij} + \left(\alpha \tau\right)_{ik\left(j\right)}
\end{align*}

* $\alpha_i$: The fixture effect

* $\beta_j$: The layout effect

* $\tau_{k\left(j\right)}: The operator random effect.

* $\left(\alpha\beta\right)$: An interaction effect between fixtures and
layouts

* $\left(\alpha\tau\right)_{ik\left(j\right)}$: A random interaction effect
between the fixture and the operator.

Since this model has both random and fixed effect terms, it is called a _mixed
effects_ model. It is fit below.

The ANOVA table suggests that layout doesn’t have much of an effect, but that
certain fixtures are easier to assemble than others. From the coefficients, it
seems that the best candidates are types 1 and 3. The fact that there is an
interaction between fixtures and operators suggests that the operators who are
much worse at some fixtures than others could be retrained.