---
title: "Split-Plot Designs"
output:
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
link-citations: yes
header-includes:
- \usepackage[fontsize=14pt]{scrextend}
---

Readings: 14.4

```{r, include=FALSE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

knitr::opts_chunk$set(message = FALSE, warning = FALSE, size = "small", echo = FALSE, fig.margin = TRUE, fig.height = 5, cache = TRUE)
```

```{r}
library("dplyr")
library("ggplot2")
library("readr")
library("EBImage")
library("reshape2")
library("tufte")
theme_set(theme_bw() + theme(legend.position = "bottom"))
```

There are times when we would like to perform a two-factor factorial experiment
across blocks, but one of the factors is much more difficult to vary than the
other. For example,

* We want to assess the effect of different irrigation (A) and corn varieties
(B) on total yield. There are three fields (Blocks) within which to gather data.
It is hard to change irrigation strategy on small subplots of land, especially
compared to corn variety.

* We run a papermill, and want to compare pulp preparation strategies (A) and
cooking temperatures (B). Samples are collected over three days (Blocks). It is
hard to change the pulp preparation strategy from sample to sample — we would
rather prepare a few big batches —  but it’s easy to cook them at different
temperatures.

This makes a true $2^{2}$ factorial experiment impractical, because it would
require randomizing over all combinations of A and B for every sample that we
collect.

* If we had 3 irrigation strategies and 6 corn varieties, we would need to
divide each field into 18 subplots and randomize the assignment of irrigation x
corn pairs.

* It’s much easier to first divide each field into 3 large plots, and then
randomly assign corn varieties to 6 subplots within each large plot.

Effectively, practical considerations impose a restriction on randomization.

## Model

The model for a split-plot design is

\begin{align*}
y_{ijk} &= \mu + \tau_{k} + \alpha_{i} + \beta_{j} + \left(\tau\alpha\right)_{ki} + \left(\alpha\beta\right)_{ij} + \epsilon_{ijk}
\end{align*}

where $\epsilon \sim N\left(0, \sigma^2\right)$ independently.

The terms can be interpreted as,

* $\mu$: The global response average.

* $\tau_{k}$: The effect of the $k^{th}$ block (e.g., $k^{th}$ field).

* $\alpha_{i}$: The effect of the $i^{th}$ level of A, the
hard-to-granularly-randomize factor (e.g., $i^{th}$ irrigation strategy).

* $\beta_{j}$: The effect of the $j^{th}$ level of B, the
easy-to-granularly-randomize factor (e.g., $j^{th}$ corn variety)

* $\left(\tau\alpha\right)_{ki}$: An interaction factor between the $k^{th}$
block and the $i^{th}$ level of $A$ (e.g., the $i^{th}$ irrigation strategy
within the $k^{th}$ field might be unusually good).

* $\left(\alpha\beta\right)_{ij}$: An interaction factor between the $i^{th}$
level of $A$ and the $j^{th}$ level of $B$.

We will typically not care about individual block effects, though we will care
about the two different treatments. Therefore, it is common to

* Use random effects for $\tau_{k}, \left(\tau\alpha\right)_{ki}$.

* Use fixed-effects for $\alpha_{i}$ and $\beta_{j}$.

The expected mean squares associated with each of the terms above can be derived
in closed form. Here, we will simply illustrate their use through the `lme4`
package. The data are from the papermill experiment described above.

```{r}
pulp <- read_csv("https://uwmadison.box.com/shared/static/843r3mda46is46nbb5b6393u7caz8orq.csv") %>%
  mutate_at(vars(Day, Method, Temperature), as.factor)
```
```{r}
ggplot(pulp) +
  geom_point(
    aes(x = Temperature, y = Strength, col = Day),
    size = 3
  ) +
  scale_color_brewer(palette = "Set2") +
  facet_grid(Method ~ .)
```

```{r}
fit <- aov(Strength ~ Method * Temperature + Error(Day / Method), data = pulp)
summary(fit)
```

## Split-Plot is not $2^{2}$ factorial

Fitting a $2^{2}$ factorial model when the data were collected with restrictions
on randomization can lead to misleading results. The code below fits an ordinary
$2^{2}$ factorial model to the papermill data. Note the overconfidence about an
effect of `Method`.

```{r}
fit <- aov(Strength ~ Method * Temperature, data = pulp)
summary(fit)
```