---
title: "Addition of Center Points"
subtitle: "Week 10 [3]"
output:
  tufte::tufte_handout:
    citation_package: natbib
    latex_engine: xelatex
link-citations: yes
header-includes:
- \usepackage[fontsize=14pt]{scrextend}
---

Reading: 6.8

```{r, include=FALSE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

knitr::opts_chunk$set(message = FALSE, warning = FALSE, size = "small", echo = FALSE, fig.margin = TRUE, fig.height = 5)
```

```{r}
library("dplyr")
library("ggplot2")
library("readr")
library("EBImage")
library("reshape2")
library("ggplot2")
library("tufte")
theme_set(theme_bw() + theme(legend.position = "bottom"))
```

```{r}
display(readImage("https://uwmadison.box.com/shared/static/n751tnavv2xgg50cj3z535t3s3dsbel0.png"))
display(readImage("https://uwmadison.box.com/shared/static/g6ge54chmmlv1guftm15fz4o4j8dz7ek.png"))
```

In section 6.7, we said that if we wanted an optimal design, we should place all our test points at the boundaries of factor’s values. This is certainly true when you have a linear factor effect. But what if the linearity assumption doesn’t hold? Which design should we pick?

```{r}
display(readImage("https://uwmadison.box.com/shared/static/6s1qw3mjzok5zrm7l384undexkb2ga2u.png"))
display(readImage("https://uwmadison.box.com/shared/static/z0rpppgsnte77f3vsl5dsrbvtyrtqne0.png"))
```

At least in the first design, we would know that something is off! We could then include polynomial terms to fit curvature in the factor effect.
This motivates the addition of _center points_. 

(pictures with center points)

Or, in the $2^2$ design,

(pictures with center points, 2^2)

## Testing

It’s possible to formally test whether there is significant quadratic curvature in the factor effect. We can add quadratic terms to our model and perform ANOVA on the associated term. Alternatively, we can look at
\begin{align}
SS_{\text{curvature}} &= \frac{n_{F}n_{C}\left(\bar{y}_{F} - \bar{y}_{C}\right)^2}{n_{F} + n_{C}}
\end{align}
```{r}
display(readImage("https://uwmadison.box.com/shared/static/w03xrmafmle36svbr4h3mpjrcjprgef7.png"))
```

where $\bar{y}_{C}$ is the average of the $n_{C}$ samples at the center point and  $\bar{y}_{F}$ is the average over all other points. A large value suggests that linearity is not plausible.

## Central Composite Design

```{r}
display(readImage("https://uwmadison.box.com/shared/static/ini2lw3vb7i89fga7ph3ywkygszaqgj6.png"))
```


If we have an unreplicated design, then we cannot fit quadratic terms — we would have more parameters than samples, and the linear system would be underdetermined. 

* A fix is to require samples at axial points.

* This design is called a _central composite_ design.