---
title: "Nonparametric ANOVA"
output: html_notebook
---

Let's generate a dataset where means are equal across all the groups. They'll be
generated from gamma distributions (normality does not hold).

```{r}
library("reshape2")
library("ggplot2")
```

```{r}
simulate_null <- function(a, n, alpha) {
  alphas <- rep(alpha, a)
  x_list <- list()
  for (i in seq_len(a)) {
    x_list[[i]] <- rgamma(n, alphas[[i]], alpha)
  }
  melt(x_list)
}
```

```{r}
x <- simulate_null(6, 50, 5)
ggplot(x) +
  geom_jitter(aes(x = L1, y = value), width=0.2)
```
There's no difference between these groups. Let's compute the nonparametric ANOVA statistic.

```{r}
kruskal.test(x$value, x$L1)$statistic
```

Let's pretend we didn't know about the theory for the reference distribution of this statistic. Let's simulate from the null many times, and see what the histogram looks like.

```{r}
sim_length <- 1000
stats <- rep(NA, ength = sim_length)

for (b in seq_len(sim_length)) {
  x <- simulate_null(6, 50, 5)
  stats[b] <- kruskal.test(x$value, x$L1)$statistic
  
  if (b %% 50 == 0) {
    p <- ggplot(data.frame(kw = stats)) +
      geom_histogram(aes(x = kw), bins=20)
    print(p)
  }
}
```


